---
subtitle: "TMA4268 Statistical Learning V2025"
title: "Compulsory exercise 1: Group XYZ (add your group number here)"
author: "NN1, NN2 and NN3 (full names of all group members)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  # html_document
  pdf_document
---
  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3,fig.align = "center")

```

```{r installation,eval=TRUE,echo=FALSE}
# install.packages("knitr")
# install.packages("rmarkdown")
# install.packages("ggplot2")
# install.packages("ggfortify")
# install.packages("MASS")
# install.packages("dplyr")
 
```

<!--  Etc (load all packages needed). -->


# Problem 1

For this problem you will need to include some LaTex code. Please install latex on your computer and then consult Compulsory1.Rmd for hints how to write formulas in LaTex.

## a)

## b)

## c)

## d)

For a MC question, just answer with the responses ordered, for example TRUE - TRUE - FALSE - FALSE (if that is the correct solution).

## e)

## f)

## g)

 

# Problem 2

Here is a code chunk (replace by what you need in the exercise):

```{r worm, eval=FALSE}
id <- "1nLen1ckdnX4P9n8ShZeU7zbXpLc7qiwt" # google file ID
d.worm <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", id))
```



## a)

## b) 
 
## c)

...


# Problem 3

# Problem 4
 